---
title: "Final Data Project"
author: "Dominique Conner"
output: html_notebook
---

# Final Data Project

## Your objectives!

To complete this project there are a few requirements you will need to fulfill. Remember that you are not on your own for this project! Data science is done best as a community, so please ask others (and instructors) questions you have when you get stuck!

1. Clearly state the data science question and goal for the analysis you are embarking on.

2. This project should be completely uploaded and up to date on GitHub. Follow the steps in `Pushing and Pulling Changes` chapter for how to git add, commit, and push the changes you have done.

3. Follow good organization principles -- you should at least have 2 folders: a `results` folder and a `data` folder. 4. 4. You should also have a README

5. Make a resulting plot that you save to a file.

6. Write up your final observations in regards to your original question. Note that some data science projects end with "This isn't what I thought it would be" or "that's strange" or "I think this is leading into another question I would need to investigate". Whatever your observations may be, write them up in your main R Markdown.

7. When you feel your analysis is ready for review, send your instructor the GitHub link to your project so they can review it.

8. Pat yourself on the back for all this work! You are a data scientist!

## Data Sources

For this project you will use whatever data you choose.

Refer back to our [Finding Data chapter](https://datatrail-jhu.github.io/DataTrail/finding-data.html) for more info on finding data.

Some options for places to find data are:

- [Kaggle](https://datatrail-jhu.github.io/DataTrail/finding-data.html#kaggle)
- [FiveThirtyEight](https://datatrail-jhu.github.io/DataTrail/finding-data.html#fivethirtyeight-data)
- You can see datasets you already have in R by running this command: `ls("package:datasets")`.

You are not limited to these options for finding your data.

<Write where you got your data and provide the link if applicable.>
#The data i used in this analysis  is from FiveThirtyEight's open data repository on Github
#Url:https://github.com/fivethirtyeight/data/blob/master/avengers/avengers.csv 
<Describe how the data was originally created. If this is data that is part of `datasets` you can use the `?` like so: ` ?datasets::AirPassengers` to see information about the datasets.Otherwise provide a summary based on the source of the data.>

## The goal of this analysis

<Write here what the goal of this analysis is. What question are we trying to answer?>
#The Goal of this analysis is to explore the Avengers dataset and identify how many heroes from the Marvel comics universe Specifically Earth-616 are full-time members of avengers team.My questions are how many times have these heroes died in the comics ? how many of the Earth-616 Avengers are females? how many of the original Avengers members are still active or alive?
## Set up

Load packages you will need for this analysis.

```{r}
library(tidyverse)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

## Set up directories

Set up the directories you will need.

```{r}
if (!dir.exists("data")) {
  dir.create("data")
}
if (!dir.exists("results")) {
  dir.create("results")
}
```


## Get the data

The dataset used in this analysis comes from FiveThirtyEight’s GitHub data repository. It contains information on 173 members of the Avengers,Here is the link:
[https://github.com/fivethirtyeight/data/raw/master/avengers/avengers.csv](https://github.com/fivethirtyeight/data/raw/master/avengers/avengers.csv)

```{r

```
```{r}
avengers<-read.csv("data/avengers.csv")
```

Explore your data here

```{r}
str(avengers)
head(avengers)
summary(avengers)
```

## Cleaning the data


```{r}
colnames(avengers)
avengers_clean<-avengers %>%filter(!is.na(Name.Alias))
avengers_clean<-avengers %>%filter(str_detect(URL,"Earth-616"))
avengers_clean<-avengers %>%filter(is.na(Full.Reserve.Avengers.Intro)|Honorary=="YES")
avengers_clean<-avengers %>%mutate(Gender==toupper(Gender))
death_columns<-avengers_clean %>%select(starts_with("Death"))
avengers_clean<-avengers_clean %>%mutate(Deaths_Total=rowSums(death_columns=="YES",na.rm=TRUE))
```

## Plot the data!

```{r}
avengers_clean %>%
group_by(Gender) %>%
summarise(Average_Deaths = mean(Deaths_Total, na.rm = TRUE)) %>%
ggplot(aes(x = Gender, y = Average_Deaths, fill = Gender)) +
geom_col() +
labs(
title = "Average Deaths by Gender (Earth-616 Avengers)",
x = "Gender",
y = "Average Number of Deaths"
) +
theme_minimal()
ggsave("results/deaths_by_gender_plot.png", width = 8, height = 6)

```

## Get the stats
```{r}
total_earth616<-nrow(avengers_clean)
females<-avengers_clean %>%filter(Gender=="FEMALE") %>%nrow()
average_deaths<-mean(avengers_clean$Deaths_Total)
most_deaths<-avengers_clean %>%filter(Deaths_Total==max(Deaths_Total)) %>%select(Name.Alias,Deaths_Total)
originals<-c("Henry Jonathan'Hank'Pym","Janet van Dyne","Anthony Edward'Tony'Stark","Robert Bruce Banner","Thor Odinson")
original_alive<-avengers_clean %>%filter(Name.Alias %in% originals)



```


### Conclusion

Write up your thoughts about this data science project here and answer the following questions:

- What did you find out in regards to your original question?From the Avengers dataset i was focused my analysis on characters from Earth-616 who were Full or Honorary Avengers.
There are over 140 Earth-616 Avengers who are either full-time or honorary members.
Approximately 13-14% of these Avengers are female, showing a historic gender imbalance on the team.Many Avengers have died in the comics on average, each character died about 0.5 to 1 time with some like Jocasta dying up to 5 times.Out of the originals Avengers(from 1963),
only Thor Odinson is still condsidered active/alive as of the data's timeline(2015).
The Rest(Iron Man,Hulk,Hank Pym,Wasp) were marked dead or inactive in some form.

- What exceptions or caveats do you have in regards to your analysis you did?
Some characters had inconsistent or missing data(e.g,,missing death records or alternate spellings). The dataset ends in 2015,so it doesn't include newer Avengers storylines or character revivals after that year. Some characters from alternate universe were present
in the data but excluded in our Earth-616 focus.





- What follow up questions do you have?
How has the Avengers' gender balance changed since 2015?
How many of the new Avengers(post-2015)are from diverse backgrounds or have major powers like reality warping?
How many total comic appearances did each member have,and is that related to how often they died?








## Print out session info

Session info is a good thing to print out at the end of your notebooks so that you (and other folks) referencing your notebooks know what software versions and libraries you used to run the notebook.

```{r}
sessionInfo()
```
